Sketch Recognition using Enhanced Particle Swarm Optimization

### 2010-12-10 Swarm/swarmModified.tex

\documentclass[a4paper, twoside]{article}

\usepackage{icgst}
%\documentclass[12pt,1p,times]{elsarticle}%[10pt]{article}
%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%\usepackage{spconf}
\usepackage{amsmath}

%\usepackage{named}
%\usepackage{hyperref}
%\usepackage[dvips,pdftex]{hyperref}
%\usepackage{subfigure}
\usepackage{array}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage[dvips,pdftex]{graphicx}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

@@

% \biboptions{}
\begin{document}

@@

%\title{Sketch Recognition using Enhanced Particle Swarm Optimization}

\twocolumn[
\scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
\title{Sketch Recognition using Enhanced Particle Swarm Optimization}
\vspace{0.1in}
\author{}  
\begin{center}
Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek abouldahab\\
\textit{Electronic Research Institute,} Cairo, Egypt,\\
\textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
{melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
%\texttt{http://www.faculty.university.edu.any}
\end{center}
\vspace{0.25in} 
]

% \twoauthors
% 
%  {Maha El Meseery, Mahmoud Fakhr El Din}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg} 
% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

@@

% \twoauthors
% {\sthanks{Corresponding Author},}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
%   {} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

%\maketitle
 \begin{abstract}
%This study

Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm. The Enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols. Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.

\end{abstract}

@@

		\includegraphics[scale=0.5]{symbolSet.PNG} 
		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
	 \end{figure}

@@

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The figures compares between the result that can be obtained using Traditional PSO by system in \cite{mypaper} and modifying the system using Enhanced PSO. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.   
    \begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment_92.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of the full feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO. The results show that Enhanced PSO increases the performance over the  Traditional PSO. }
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment_68.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of half the feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO. The results show that Enhanced PSO increases the recognition rate. }
   \label{exp1}
\end{figure}

@@

@@

\bibliographystyle{plain}
\bibliography{../../../neededfiles/Bibliographies/Mybibliography}

>>>>

%\documentclass[a4paper, twoside]{article}

%%\usepackage{icgst}
%\documentclass[12pt,times,a4paper]{article}%[10pt]{article}
%% Use the option review to obtain double line spacing
 \documentclass[preprint,1p,times,review]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%\usepackage{spconf}
\usepackage{amsmath}
%\usepackage{named}
%\usepackage{hyperref}
%\usepackage[dvips,pdftex]{hyperref}
%\usepackage{subfigure}
\usepackage{array}
\usepackage{amssymb}
%\usepackage{epsfig}
%\usepackage{epstopdf}
%\usepackage{xltxtra}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{sectsty}
%\sectionfont{\normalsize  }
%\subsectionfont{ \normalsize\it \textbf }
%\subsubsectionfont{\normalsize \it \textbf }
%\paragraphfont{\large\it}
%\subparagraphfont{\normalsize\it}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

@@

% \biboptions{}
% \setlength\hoffset{0in}
% % \setlength\topmargin{0in}
% % \setlength\headheight{0in}
% % \setlength\headsep{-0.5in}
% % \setlength\textheight{10.0in}
%  \setlength\textwidth{7in}
%  \setlength\oddsidemargin{-0.5in}
%  \setlength\evensidemargin{-0.5in}
% % \setlength\parindent{0.25in}
% % \setlength\parskip{0.25in}
% \linespread{1.5}
\begin{abstract}

%This study
%%%This study

Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Algorithms to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. A comparision is presented between using Traditional Particle Swarm Optimization and the Enhanced Particle Swarm Optimization (EPSO) which is based on the perturbation of the particle velocity. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS). Results also show that using an Enhancment on the Tradiontal Swarm algorithm improves recognition rate.

\end{abstract}
\begin{document}

@@

\title{Sketch Recognition using Enhanced Particle Swarm Optimization}

% \twocolumn[
% \scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
% \title{Sketch Recognition using Enhanced Particle Swarm Optimization}
% \vspace{0.1in}
% \author{}
% \begin{center}
% Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek
%abouldahab\\
% \textit{Electronic Research Institute,} Cairo, Egypt,\\
% \textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
% {melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
% %\texttt{http://www.faculty.university.edu.any}
% \end{center}
% \vspace{0.25in}
% ]

% \twoauthors
%
%  {Maha El Meseery, Mahmoud Fakhr El Din}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company
%\\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

@@

\author{
\small{
Maha El Meseery \footnote{Corresponding Author} \textsuperscript{a}, Mahmoud
Fakhr El Din \textsuperscript{a}, Tarek abouldahab \textsuperscript{b}}\\
 \textsuperscript{a} \small{Signals Processing Group, Computers and Systems
Department,}\\ \small{ Electronic Research Institute, Cairo, Egypt.}\\
  \small {\textsuperscript{b}Cairo Metro Company ,
   Ministry of transport , Cairo, Egypt.}\\
   \small{melemseery@eri.sci.eg,  mafakhr@mcit.gov.eg, heshoaboda@hotmail.com
   }
}
\date{}
% {\sthanks{Corresponding Author},}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
%   {Tarek abouldahab} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
%  {Maha El Meseery \sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company
%\\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

\maketitle
%  \begin{abstract}
%
% %This study
% %%%This study
%
% Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
%
%
%  \textsl{Sketches and drawings are widely used as a simple method of expressing
% thought and designs. The goal of this paper is to improve our sketch recognition
% system by using an enhanced particle swarm optimization algorithm. The Enhanced
% Particle Swarm Optimization (EPSO) is based on the perturbation of the particle
% velocity is used to optimally segment the strokes the user draws into meaningful
% geometric primitives. These geometric primitives are grouped to formulate
% symbols which are further identified using a Support Vector Machines (SVM)
% classifier. Experiments were conducted on the benchmark dataset of simple
% presentation (Hs-DS) symbols. Results show that the enhanced particle swarm
% optimization improves the final recognition accuracy.}
% \end{abstract}

@@

\fbox{ \parbox{5cm}{%
		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
		\caption{The Hs-DS Symbol Set} The dataset consist of simple presentation symbols.
		\label{fig:HsSet}
\end{figure}
\begin{table}
\begin{center}
\scalebox{0.8}{
 \begin{tabular}{|p{3cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
 \hline
 & Hs-DS &  EL-DS & LD-DS \\ \hline
No of samples  &  7791 & 2764 & 1859 \\ \hline
No of categories  & 13 & 15 & 8 \\ \hline
Avg. Samples per Category& 600 & 184 & 232 \\\hline
No of users  & 16 & 7 & 7 \\  \hline
Balanced & Yes&No & No \\ \hline
Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
Examples & Figure \ref{fig:HsSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
\end{tabular}
}
\caption[Datasets Comparisons]{Table compares between datasets used in experiments in terms of number of samples, number of samples per category, etc...}
\label{tab:datasets}

\end{center}
\end{table}
\begin{figure}
\centering

\fbox{ \parbox{5cm}{%
		\includegraphics[scale=0.5]{images/symbolSet.PNG}}}
		\caption{The Hs-DS Symbol Set} The dataset consist of simple
presentation symbols.
		\label{fig:HsSet}
\end{figure}
 \begin{figure}
			\subfigure[The Digital Design Symbols Dataset(LD-DS)]{\label{fig:LogicsymbolSet} \includegraphics[scale=0.5]{images/logicSet.jpg}}
			\hfill
			\subfigure[The Electrical Symbols Dataset(EL-DS)]
			 {\label{fig:ELsymbolSet}\includegraphics[scale=0.45]{images/EelectImage.jpg}}

	\caption{The datasets that were manually collected}  a)  The (LD-DS) dataset consist of symbols that are visually similar (i.e. has low shape diversity). b) The (EL-DS) dataset consist of complex topological symbols.
\end{figure}
 \begin{figure}\centering
			%\subfigure[The Digital Design Symbols Dataset(LD-DS)]{

\includegraphics[scale=0.5]{images/logicSet.jpg}\label{fig:LogicsymbolSet}%}
	\caption{The Digital Design Symbols Dataset(LD-DS)}    The (LD-DS)
dataset consist of symbols that are visually similar (i.e. has low shape
diversity).
\end{figure}
\subsection{Segmentation Algorithms comparisons}
\label{sec:AlgExp}
We performed several experiments to evaluate the presented recognition system. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} and \ref{fig:testLD} show the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. Combining the ellipse detection module improves the performance in certain datasets. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.  The results also shows that combining (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforms AlgS2 in Hs-Ds only. This behavior is explained by noticing that features are computed based on the chosen segmentation. This means that different segmentation for a symbol generates a different set of features used for recognition. This behavior affects Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure nature of the symbols in both datasets than of Hs-Ds.

 \begin{figure}
	\centering
	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
 		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
 		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgEL.jpg}}
	 	\caption{ The recognition rate of different algorithms. }
 \end{figure}
\subsection{Swarm Algorithms Experiments}
Another set of tests were performed on the system to enhance non-ellipse fitting
algorithms. The effect of the number of particles, the error thresholds and
various other parameters were investigated. The results of these tests are shown
in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure
\ref{fig:swarmtesting} shows that as the number of swarm iterations increase the
error value decrease (error of the segmentation algorithm). The final number of
particles and maximum swarm iterations used in the system was based on a
tradeoff between error achieved versus the computation time. The effect of other
algorithm parameters such as $c_1$, $\omega$ (mentioned in
Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure
\ref{fig:swarmtesting2},\ref{fig:swarmwtest}).%, $c_2$

 \begin{figure}
	\centering
	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{
\label{fig:swarmtesting}
\includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}

		\subfigure[Segmentation Error Vs. $c_1$]{
\label{fig:swarmtesting2}  \includegraphics[scale=0.3]{images/C1Swarm.jpg}}
%swarmParamterC.jpg

				\subfigure[Segmentation Error Vs.
$\omega$]{\label{fig:swarmwtest}
\includegraphics[scale=0.3]{images/wSwarm.jpg}} %swarmParamterC.jpg

	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error
decrease as the number of swarm iteration increase until it reach a saturation.
 b) The behavior of segmentation error versus $c_1$. c) The behavior of
segmentation error versus  $\omega$.

\end{figure}

%In our experiment we used the same benchmark dataset used in \cite{mypaper},
%the dataset was collected by Hse and Newton\cite{HeloiseBeautification}. The
%data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to
%50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data
%set. Five different splits were generated to divide the dataset into training
%and test sets. The results displayed are the average recognition accuracy of the
%five splits. The accuracy is computed as the number of correctly recognized
%samples divided by the total number of samples in the test.

%  \begin{figure}
%   \centering
%
% 		\includegraphics[scale=0.5]{symbolSet.PNG}
% 		\caption[The Symbol Set] {The symbol set}
%\label{fig:symbolSet}
% 	 \end{figure}

@@

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The figures compares between the result that can be obtained using Traditional PSO and modifying the system using Enhanced PSO. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.
    \begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment_92.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of the full feature set on both
The traditional PSO system and the Enhanced PSO. The results show that Enhanced PSO increases the performance over the  Traditional PSO. }
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment_68.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of half the feature set on both
the traditional PSO system and Enhanced PSO. The esults show that Enhanced PSO increases the recognition rate. }
   \label{exp1}
\end{figure}

@@

This paper presented a new approach to sketch recognition using DPSO. The system
uses multiple preliminary information, i.e., speed and curvature information, as
input to the \textit{DPSO}. Results show that the \textit{DPSO} in general
generates an optimized stroke segmentation which improves the final recognition
rate.  The final recognition is based on both global shape and hierarchical
stroke properties. Adding stroke and segment geometrical features improved the
final recognition. Global shape and moment properties were used to prevent the
need to use graph matching which is computationally expensive
\cite{SketchRead2007}. The experiments evaluated two \textit{DPSO} different
segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds,
EL-Ds and LD-Ds). Results show that AlgS2 achieves better segmentation and final
accuracy than Alg3 \cite{earlyprocess} on the datasets used. The results proved
that the system is easily expandable to more and more symbols as it does not
depend on low level primitive recognizers but on high level symbol and
geometrical features.  The system achieved an average overall improvement of
more than 2\% over Alg3 in all datasets.

  The tradeoff between accuracy achieved and time complexity can be further
investigated to achieve better results. Currently, the system processes one
symbol at a time, an enhancement would be to allow users to draw more than one
symbol at a time and incorporate a method for separating symbols. Thus, possible
extension of this research is to complete the clustering algorithm for fully
automated sketch recognition. Another area of enhancements is the features
extraction methods. Introducing more spatial and geometrical features is
believed to improve classifications. Features that represent the appearance of
the symbols are likely to improve the recognition of symbols that have similar
geometrical structure (i.e. Nor and OR gate in logic symbols).

@@

%\bibliographystyle{natbib}
\bibliographystyle{ieeetr}
\bibliography{../../../neededfiles/Bibliographies/Mybibliography}
